---
title: "Reproducible Research: Peer Assessment 1"
author: "Gerardo Santiago Flores"
date: "June 12, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(plyr)
library(ggplot2)
```

## Loading and preprocessing the data
### 1. Code for reading in the dataset and/or processing the data
```{r read_data}
read_data <- read.csv("activity.csv", header = TRUE, na.strings = "NA")
head(read_data)
```


## What is mean total number of steps taken per day?
### 2. Histogram of the total number of steps taken each day
```{r hist_total_per_day}
total_steps <-ddply(read_data, .(date), summarise, sum=sum(steps))
hist(total_steps$sum, xlab = "Steps", main ="Total number of steps per day ")
```

## What is the average daily activity pattern?
### 3. Mean and median number of steps taken each day
```{r mean_meadian_each_day}
mean_median_data <- ddply(read_data, .(date), summarise, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))
mean_median_data

```

### 4. Time series plot of the average number of steps taken

```{r plot_average_steps}
average_steps_per_interval <- ddply(read_data, .(interval), summarise, mean=mean(steps, na.rm = TRUE))
plot(average_steps_per_interval, type = "l", main = "Mean number of steps per interval")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r max_steps_interval}
interval_max_steps <- average_steps_per_interval$interval[which.max(average_steps_per_interval$mean)]
print("interval with max steps")
print(interval_max_steps)
```

## Imputing missing values
### 6. Code to describe and show a strategy for imputing missing data
The strategy used is to replace NA by mean value of that 5-minute interval

```{r total_na}
print("Total number of rows with NA")
print(sum(is.na(read_data$steps)))
```

```{r fill_stategy}
print("Fill NA strategy")
intervals <- unique(read_data$interval)
filled_data <- read_data
for (inter in intervals)
{
    index_to_fill <- which(is.na(read_data$steps) & read_data$interval == inter)
    filled_data[index_to_fill,]$steps <- average_steps_per_interval[average_steps_per_interval$interval == inter,]$mean
}
print(head(filled_data))
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r hist_total_per_day_filled}
filled_total_steps <-ddply(filled_data, .(date), summarise, sum=sum(steps))
hist(filled_total_steps$sum, xlab = "Steps", main ="Total number of steps per day (Filled Data)")
```

```{r mean_meadian_each_day_filled}

filled_mean_median_data <- ddply(filled_data, .(date), summarise, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))
print("Mean Meadian of filled data")
print(filled_mean_median_data)

```

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays_mean}
filled_data$date <- as.Date(filled_data$date)
filled_data$weekday <- isWeekday(filled_data$date)
weekday_average_steps_per_interval <- ddply(filled_data, .(interval, weekday), summarise, mean=mean(steps, na.rm = TRUE))
ggplot(weekday_average_steps_per_interval, aes(x = interval, y = mean, colour = weekday)) + geom_line() + facet_grid(weekday ~ .) + ggtitle("Mean steps per interval") + labs(x="Interval",y="Number of steps") 
```